\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-07-10 ResponseToReview.tex

4) Nevertheless, the manuscript is worth publishing as it does present data that may help rule out some possible explanations as to why tb1 is so common in parviglumus.\\

\res{Thank you; we hope that our paper in conjunction with future research and experiments can help to further the understanding of the role of both \emph{tb1} and the \emph{Hopscotch} insertion in teosinte.}

5) This paper is fairly maize-centric. A more general message is included in the discussion, not as much in the introduction, and not at all in the abstract. This could be easily remedied.\\

\jri{tweak intro and abstract, add a bit to discussion maybe, explain text changes here.}
\lev{Jeff & Matt, this is already in the intro. I don't see what more we can say that would be appropriate. ``The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte."}

@@

\res{Thank you.}.

Specific suggestions follow:

7) Introduction:
Page 4, Lines 7 - 9: You say "Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic eﬀects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".  Can you tell us why an understanding is useful? Can you give some examples?\\

\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice (Kovach and McCouch 2008).}

8) Page 5, line 12 - 14: You say "The tb1 locus appears to play an important role in the shade avoidance pathway in Zea mays and other grasses and may therefore be crucial to the ecology of teosinte (Kebrom and Brutnell, 2007; Lukens and Doebley, 1999)." Please clarify; do the references belong to the first half of the sentence, supporting tb1 playing a role in Zea mays, or the second half of the sentence, supporting tb1 possibly being crucial to the ecology of teosinte? Which of the two distinct things have actually been studied? If the first, could you tell us how tb1 has been shown to work in maize under shade conditions? If the second, was it actually studied, or merely hypothesized in teosinte?\\

@@

\res{We have clarified issues regarding the references for this section of the paper.}

9) Page 5, Lines 15 - 16: you say you ``aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic eﬀects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed. Did you phenotype mexicana? Can you present the data here, if so? Does the analysis of this group of teosintes add anything to your conclusions?\\

@@

\res{Reviewer 2 is correct in noting that ssp. \emph{mexicana} is not as prevalent in our paper as ssp. \emph{parviglumis}. This fact is due to limitations in access to accessions of ssp. \emph{mexicana} as field collections for ssp. \emph{mexicana} are not nearly as extensive as those of ssp. \emph{parviglumis}. Additionally ssp. \emph{mexicana} are more difficult to cultivate in a greenhouse setting. We felt that it would be important to include ssp. \emph{mexicana} in our study, especially since it is known that gene flow occurs between ssp. \emph{mexicana} and landrace maize. We have noted in our methods that we were only able to phenotype a population from which we had a large number of seed collections.}

10) Materials and Methods:
Page 5, lines 22 - 23: how many individuals per accession? A fixed number, or variable, and depending on what? This can be mentioned so the reader does not have to go to the appendix or supplements.\\

\res{We have added a range of values (1-38 individuals) per population in the methods: ``We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article)."}

11) Page 5, Line 28: mention how big the hopscotch element is.  Can it be amplified in one amplicon? Again, this would be useful to know without having to look in appendix or supplements.\\

### 2014-07-10 manuscript.tex
edits
edits

\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article). DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}. We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion. Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR). Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'}; 
HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article). Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $\sim$1.1kb), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article). When only one PCR resolved well, we scored one allele for the individual. We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min. PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.

>>>>

\textbf{Sampling and genotyping---}We sampled 1,110 individuals from 350 accessions (247 maize landraces, 17 \emph{mexicana} populations, and 86 \emph{parviglumis} populations; ranging from 1-38 individuals per population) and assessed the presence or absence of the \emph{Hopscotch} insertion (Appendix 1 and Appendix 2, See Supplemental Materials with the online version of this article). DNA was extracted from leaf tissue using a modified CTAB approach \citep{DoyleDoyle1990, Maloof1984}. We designed primers using PRIMER3 \citep{RozenSkaletsky2000} implemented in Geneious \citep{Kearse2012} to amplify the entire \emph{Hopscotch} element, as well as an internal primer allowing us to simultaneously check for possible PCR bias between presence and absence of the \emph{Hopscotch} insertion. Two PCRs were performed for each individual, one with primers flanking the \emph{Hopscotch} (HopF/HopR) and one with a flanking primer and an internal primer (HopF/HopIntR). Primer sequences are HopF, {\small 5'-TCGTTGATGCTTTGATGGATGG-3'}; 
HopR, {\small 5'-AACAGTATGATTTCATGGGACCG-3'}; and HopIntR, {\small  5'-CCTCCACCTCTCATGAGATCC-3'} (Appendix 3 and Appendix 4, See Supplemental Materials with the online version of this article). Homozygotes show a single band for absence of the element ($\sim$300bp) and two bands for presence of the element ($\sim$5kb and $\sim$1.1kb), whereas heterozygotes are three-banded (Appendix 2, See Supplemental Materials with the online version of this article). When only one PCR resolved well, we scored one allele for the individual. We used Phusion High Fidelity Enzyme (Thermo Fisher Scientific Inc., Waltham, Massachusetts, USA) and the following conditions for amplifications: 98$^{\circ}$C for 3 min, 30 cycles of 98$^{\circ}$C for 15 s, 65$^{\circ}$C for 30 s, and 72$^{\circ}$C for 3 min 30 s, with a final extension of 72$^{\circ}$C for 10 min. PCR products were visualized on a 1\% agarose gel and scored for presence/absence of the \emph{Hopscotch} based on band size.
